---
title: "Biodiversity, function, and soil "
author: "Emily P. Piché"
date: "9/12/2018"
output: 
  beamer_presentation: 
    colortheme: dolphin
    highlight: espresso
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


--- CURRENTLY INCOMPLETE ---

## Week 3 assigment instructions
https://gotellilab.github.io/Bio381/Homeworks/Homework03_F2018.html


Create a short Beamer presentation (6 to 12 slides). Note that you will need to save your images to .tiff. or .jpeg or other digital formats and put them in a folder that can be linked to from your Beamer code.

Spend a bit of time exploring the different themes, colors, and fonts available in Beamer, and set up something that looks nice to you.

Your presentation should include the following:

- \color{gray}sequential bullet points
- simultaneous bullet points
- colored text
- images
- images with captions
- images with a caption overlaid (you may need to change the text color for better contrast)


## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## Slide with one image



## Slide with two images

\begin{picture}(320,250)
\put(0,120){\includegraphics[height=4cm]{IMG_1407.JPG}}
\put(180,120){\includegraphics[height=4cm]{IMG_1407.JPG}}
\put(10,75){\begin{minipage}[t]{1\linewidth}
  {This is a picture of soils from \textit{Mt Mansfield} -
  \textbf{I set them on fire first}. They're colorful!}
  \end{minipage}}
  \end{picture}

