---
title: "Homework 6.2"
author: "Emily P. Piché"
date: "10/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
### Read in your own data file
z <- read.csv("/Users/emilypiche/Desktop/soildata.csv", as.is=TRUE,header=TRUE,sep=",",stringsAsFactors=FALSE)
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

Create a distribution to use for the following chunks. After running them, go back and re-run your own data through.
```{r}
# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,]
```

Plotting a histogram of the data:
```{r}
# made a vector of the LOI column called "SOM"
SOM <- z$LOI
p1 <- ggplot(data=z, aes(x=SOM, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk") 
print(p1)
```

Add empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

Get maximum likelihood parameters for normal
```{r}
normPars <- fitdistr(SOM,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] 
normPars$estimate["sd"] 
```

Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$LOI),len=length(SOM))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(SOM), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

Plot exponential probability density
```{r}
expoPars <- fitdistr(SOM,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(SOM), args = list(rate=rateML))
 p1 + stat + stat2
```
 
Plot uniform probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(SOM), args = list(min=min(SOM), max=max(SOM)))
 p1 + stat + stat2 + stat3
```


Plot gamma probability density
```{r}
gammaPars <- fitdistr(SOM,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(SOM), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```
 
Plot beta probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=SOM/(max(SOM + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=SOM/max(SOM + 0.1), start=list(shape1=1,shape2=2), "beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

shape2ML
betaPars

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(SOM), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

summary(SOM)

```

Creating a gamma distribution with the features of my own data
```{r}
gammaPars <- fitdistr(SOM,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

myGamma <- rgamma(n=length(SOM), shape=shapeML, rate=rateML)
gammaDist <- qplot(x=myGamma, color=I("black"),fill=I("goldenrod"))

xval <- seq(0,max(myGamma),len=length(myGamma))

fakeGamma <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(SOM), args = list(shape=shapeML, rate=rateML))

gammaDist + fakeGamma

```

I found that though the gamma distribution seemed to fit my soil organic matter data best. But, when I simulated the data and fit the gamma probability density curve, I found that the curve did not seem to fit the simulated data. Instead, it seems to show much lower values than the simulation actually produced.

Here are the summaries of my real data versus the simulated data:
```{r}
summary(SOM)
summary(myGamma)
```

