---
title: "Homework 7"
author: "Emily P. Piché"
date: "10/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

A pattern I hypothesized that I would see at my sites is a positive relationship between soil organic matter (%) and the proportion of conifers on a plot (%). After collecting my data this summer, I found that at my 23 sites, the proportion of conifers on each plot explained 25% of the variation in soil organic matter. 

Here I will create a fake dataset to explore this hypothesis in R using a normal distribution. 


Sample sizes and means of my soil organic matter and conifer data:
```{r}
# calling up MASS and ggplot2
library(MASS)
library(ggplot2)
library(dplyr)
# reading in my own data from a csv
PL <- read.csv("/Users/emilypiche/Desktop/plotlevel.csv", as.is=TRUE)
#SOM - real soil organic matter data
SOM <- PL$LOI
#RC - real conifer proportion data
RC <- PL$CONIFER_D

# SOM parameters
normPars1 <- fitdistr(SOM,"normal")
normPars1
# conifer parameters
normPars2 <- fitdistr(RC,"normal")
normPars2

```

Assuming these two groups follow a normal distribution (propotion of conifers in each plot, and soil organic matter content (%)), I simulated the following distributions based on parameters from my actual data. I had 23 sites, the mean proportion of conifers was .33, and the standard deviation was 0.325. I created the fake SOM data from the fake conifer data (rather than generating it on its own) because I wanted to generate data that would show a positive relationship with a simple regression, and did not use the actual parameters from my SOM data (n=23, mean=.33, sd=.20)
```{r}
fCon <- rnorm(length(RC), normPars2$estimate["mean"], normPars2$estimate["sd"]) # creating fake conifer proportions
e <- rnorm(23, .2, .1) # creating some noise

# OM content is a function of the proportion of conifers on each plot, so simulated SOM is created using a noise factor and the fCon data
fSOM <- 0 + .1*fCon+ e 

# putting data into a dataframe 
simData <- data.frame(fCon, fSOM)
head(simData)
```

The following chunk shows the results of a linear model. I found on the first run that change in generated proportion of conifers on each plot explained 34% of the variation in generated soil organic matter content, which is a bit higher than I what was attempting to achieve.

```{r}
# summary of a linear model of my generated conifer and SOM data
sum <- summary(lm(simData$fSOM~simData$fCon))
r2<-sum$adj.r.squared
p<-sum$coefficients[7]
# r^2 of the linear model
r2
# p value of linear model
p
```

Here is figure showing one version of the generated data:
(This may or may not show a significant relationship depending on the generated data)
```{r}
qplot(simData$fCon,simData$fSOM,
      xlab="Simulated proportion of Conifers",
      ylab="Simulated soil organic matter (%)")+ theme_bw() + geom_smooth(method="lm")
```

The r-squared varied quite a bit each time I ran the code, which is likely a factor of the low sample size (.05 - .45). In my real dataset, each site is represented with 3 different cores which are averaged for each site. I think that if I generated the code in this way, it would reduce some of the variation between runs. I also found that I could not constrain the proportion of conifers in the overstory to only include positive numbers because I am doing this exercise using a normal distribution.

In this chunk below, I have adjusted the level of noise in my SOM data to see how much it reduces the variability between runs through the linear model. I found that each time I created new SOM data, it showed a significant relationship.
```{r}
e2 <- rnorm(23, .5, .05) # creating less noise with 0.05 sd rather than 0.5, and increasing the mean to .5

# new fake SOM data
fSOM2 <- 0 + .1*fCon+ e2 

# new dataframe with fSOM2
simData2 <- data.frame(fCon, fSOM2)
head(simData2)

# summary of a linear model of my generated conifer and SOM data
sum2 <- summary(lm(simData2$fSOM2~simData$fCon))
unlist(sum2)
R2<-sum2$adj.r.squared
P<-sum2$coefficients[7]
# r^2 of the linear model
R2
# p value of linear model
P

# plot of data with SOM that has less generated noise
qplot(simData2$fCon, simData2$fSOM2,
      xlab="Simulated proportion of Conifers",
      ylab="Simulated soil organic matter (%)")+ theme_bw() + geom_smooth(method="lm")
```

